### 2024-01-17 parts/correction.tex
residual transducer??

\begin{remark}
    \label{inductive-correct-poly:remark}
    Let $P \in \Rel[X,\vec{Y}]$ be a polynomial.
    The following are equivalent:
    \begin{enumerate}
        \item $P \in \CorrectPoly$,
        \item The \kl{maximal monomials} of 
            $P$ are \kl{non-negative}
            and 
            for all $n \in \Nat$,
            $\restr{P}{X = n} \in \CorrectPoly$.
    \end{enumerate}
    Furthermore, if $P \in \Rel[]$ is a polynomial without
    indeterminate, then $P \in \CorrectPoly$ if and only
    if its \kl{maximal monomial} (the constant term)
    is \kl{non-negative}.
\end{remark}
\begin{proof}
    By induction on the number of indeterminate,
    leveraging the fact that if there exists at least one indeterminate
    $X$, i.e., that $\vec{X} = X \vec{Y}$, then for all $n \in \Nat$
    and partial function $\nu \colon \vec{Y} \topartial \Nat$,
    we have 
    \begin{equation*}
        \restr{\left(\restr{P}{X = n}\right)}{\nu} = \restr{P}{ \nu[X = n]}
        \quad
        ,
    \end{equation*}
    where $\nu[X = n]$ is the partial function from
    $\vec{X}$ to $\Nat$ that coincides with $\nu$ on $\vec{Y}$
    and maps $X$ to $n$.
\end{proof}

@@

\begin{lemma}
    \label{derivation-translation:lem}
    Let $P \in \Rel[\vec{X}]$, the following are equivalent
    \begin{enumerate}
        \item $P \in \CorrectPoly$,
        \item $P$ is \kl{non-negative}
            and there exists a computable $K \in \Nat$
            such that $\translate{K}(P) \in \Nat[\vec{X}]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    \textbf{TODO: do the proof}.
    To prove \cref{corrected-5:item} $\implies$ \cref{corrected-1:item}, it
    suffices to remark that for every polynomial $P$, for every partial
    valuation $\nu \colon \vec{X} \topartial \Nat$, for every $K \in \Nat$, the
    \kl{maximal monomials} of $\restr{\translate{K}(P)}{\nu}$ are exactly the
    \kl{maximal monomials} of $\restr{P}{\nu}$, the former being
    \kl{non-negative} if $\translate{K}(P) \in \Nat[\vec{X}]$.

    Conversely, the implication \cref{corrected-1:item} $\implies$
    \cref{corrected-5:item} follows immediately from the fact that polynomials
    in $\CorrectPoly$ are \kl{non-negative}, and on the following induction.
    Let $P \in \CorrectPoly$, if $P$ is constant, then $\translate{0}(P) \in \Nat[\vec{X}]$.
    Otherwise, if $P = P_1 + P_2$ where $P_1$ is the sum of the \kl{maximal
    monomials} of $P$, there exists $K \in \Nat$ such that $Q \defined
    \Diff{K}{P_1}+ P_2$ belongs to $\CorrectPoly$,
    and by induction hypothesis there exists $L \in
    \Nat$ such that $\translate{L}{Q} \in \Nat[\vec{X}]$. As a consequence,
    $\translate{K+L}(P) = \translate{L}(P_1) + \translate{K}(\translate{L}(Q))$
    is a sum of two polynomials in $\Nat[\vec{X}]$, and we have concluded.
\end{proof}

@@

While \cref{lem:correct-to-n-rat} provides effective conversion, it does not
explicitly state that the nembership is decidable. However, it is not too
difficult to remark that $\CorrectPoly$, the proof scheme can be followed to
obtain decidability. For the sake of completeness, we provide a self-contained
proof hereafter, that furthermore

@@

\begin{theorem}
    \label{corrected-version:thm}
    Let $P \in \Rel[\vec{X}]$ be a polynomial.
    The following are equivalent:
    \begin{enumerate}
        \item \label{corrected-1:item} $P \in \CorrectPoly$,
        \item \label{corrected-2:item} $P$ is \kl{represented} by a \kl{$\Nat$-rational series},
        \item \label{corrected-3:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
        \item \label{corrected-4:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
        \item \label{corrected-5:item} $P$ is \kl{non-negative}
            and there exists $K \in \Nat$ such that $\translate{K}(P) \in \Nat[\vec{X}]$.
    \end{enumerate}
    Furthermore, the membership is decidable, and effective conversion
    procedures exist between all the representations.
\end{theorem}
\begin{proof}
    The implications 
    \cref{corrected-4:item} $\implies$
    \cref{corrected-3:item} $\implies$
    \cref{corrected-2:item} are obvious.
    \cref{lem:correct-to-n-rat} proves
    \cref{corrected-1:item} $\implies$ \cref{corrected-4:item},
    while \cref{n-rat-correct:lem}
    proves 
    \cref{corrected-2:item} $\implies$ \cref{corrected-1:item}.
    Note that the lemmas provide effective conversion procedures,
    and that \cref{lem:correct-to-n-rat} also provides a decision
    procedure.

    The equivalence between
    \cref{corrected-5:item} and \cref{corrected-1:item}
    is precisely \cref{derivation-translation:lem}.
\end{proof}

@@

>>>>

@@

@@

While \cref{lem:correct-to-n-rat} provides effective conversion, it does not
explicitly state that the membership is decidable. However, it is not too
difficult to remark that $\CorrectPoly$, the proof scheme can be followed to
obtain decidability. For the sake of completeness and readability,
we extract this fact in the following lemma.

@@

\begin{theorem}
    \label{corrected-version:thm}
    Let $P \in \Rel[\vec{X}]$ be a polynomial.
    The following are equivalent:
    \begin{enumerate}
        \item \label{corrected-1:item} $P \in \CorrectPoly$,
        \item \label{corrected-2:item} $P$ is \kl{represented} by a \kl{$\Nat$-rational series},
        \item \label{corrected-3:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
        \item \label{corrected-4:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
    \end{enumerate}
    Furthermore, the membership is decidable, and effective conversion
    procedures exist between all the representations.
\end{theorem}
\begin{proof}
    The implications 
    \cref{corrected-4:item} $\implies$
    \cref{corrected-3:item} $\implies$
    \cref{corrected-2:item} are obvious.
    \cref{lem:correct-to-n-rat} proves
    \cref{corrected-1:item} $\implies$ \cref{corrected-4:item},
    while \cref{n-rat-correct:lem}
    proves 
    \cref{corrected-2:item} $\implies$ \cref{corrected-1:item}.
    Note that the lemmas provide effective conversion procedures,
    and that \cref{lem:correct-to-n-rat} also provides a decision
    procedure.
\end{proof}

@@

It was conjectured in the 
\cite[Open question 5.55]{gaetanphd}.
\cite[Conjecture 7.61]{gaetanphd}
Let us remark that a previous conjecture was not true, and provide the
recursively defined variant. This can be understood in the light of
\cref{corrected-version:thm}, as the fact that \kl{$\Nat$-rational polynomials}
are essentially $\Nat$-polynomials.

\begin{example}
    The polynomial $\BadPoly$ is not a 
    \kl{$\Nat$-rational polynomial},
    but is \kl{non-negative} and satisfies
    $\translate{10}(\BadPoly) \in \Nat[\vec{X}]$.
\end{example}

\begin{lemma}
    \label{derivation-translation:lem}
    Let $P \in \Rel[\vec{X}]$, the following are equivalent
    \begin{enumerate}
        \item \label{d-t-correct:item} $P \in \CorrectPoly$,
        \item \label{d-t-transl:item}
            There exists $K \in \Nat$,
            such that for every partial
            function $\nu \colon \vec{X} \topartial \Nat$,
            $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    To prove \cref{d-t-transl:item} $\implies$
    \cref{d-t-correct:item}, let us remark that
    if $P$ is \kl{non-negative} and $\translate{K}(P) \in \Nat[\vec{X}]$,
    then $\translate{K}(P) \in \CorrectPoly$ thanks to
    \cref{n-poly-n-poly:example}.
    In particular, 
    then for every partial valuation 
    $\nu \colon \vec{X} \topartial \Nat$, 
    $\restr{\translate{K}(P)}{\nu} \in \Nat[\vec{X}]$,
    and in particular conclude that
    \kl{maximal monomials} of $\restr{\translate{K}(P)}{\nu}$ are exactly the
    \kl{maximal monomials} of $\restr{P}{\nu}$, the former being
    \kl{non-negative} if $\translate{K}(P) \in \Nat[\vec{X}]$.

    Conversely, the implication \cref{d-t-correct:item} $\implies$
    \cref{d-t-transl:item} is proven by induction on $\MaximalMonomials(P)$.
    If $P$ is a constant polynomial, then $P = n$ for some $n \in \Nat$,
    and we conclude that $\translate{0}(P) \in \Nat[\vec{X}]$.
    Otherwise, $P = P_1 + P_2$ where $P_1$ is the sum of the \kl{maximal
    monomials} of $P$.
    Using \cref{derivation-stabilises-correct:lem},
    there exists a computable $K \in \Nat$
    such that 
    $Q \defined
    \Diff{K}{P_1} + P_2$ belongs to $\CorrectPoly$.
    Since, $\MaximalMonomials(Q) \hoarele \MaximalMonomials(P)$,
    we conclude by induction hypothesis that
    there exists $L \in \Nat$ such that 
    for every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
    $\translate{L}(\restr{Q}{\nu}) \in \Nat[\vec{X}]$. 
    Similarly,
    for all partial functions $\nu \colon \vec{X} \topartial \Nat$
    fixing at least one indeterminate (appearing in $P$),
    $\MaximalMonomials(\restr{P}{\nu}) \hoarele \MaximalMonomials(P)$,
    hence there exists $L_\nu \in \Nat$ such that
    $\translate{L_\nu}{\restr{P}{\nu}} \in \Nat[\vec{X}]$.
    Let us define $L_m$ to be the maximum of $L$, and $L_\nu$
    where $\nu$ ranges over partial functions described above.

    Let us prove that for all partial functions
    $\nu \colon \vec{X} \topartial \Nat$,
    $\translate{L_m + K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
    If $\nu$ is the empty valuation, then $\restr{P}{\nu} = P$,
    and
    $\translate{L_m + K}(P) = \translate{L_m}(P_1) + \translate{K+L_m}(Q)$
    is a sum of two polynomials in $\Nat[\vec{X}]$.
    Otherwise,
    $\translate{L_m + K}(\restr{P}{\nu}) = \translate{K}(\translate{L_m}(\restr{P}{\nu}))$
    belongs to $\Nat[\vec{X}]$ since the latter is stable under
    positive translations.
\end{proof}

### 2024-01-17 parts/prelim.tex
residual transducer??

@@

>>>>

Remark II.21. It follows from Remark II.6, [11, Proposition 6.1], and Theorem
II.20, that Z-rational series of polynomial growth are exactly those computable
by weigthed automata with coefficients in {0,1,−1} of polynomial ambiguity. We
are not aware of a direct proof of this correspondence.

@@

Let us briefly state that one can decide whether a \kl{$\Rel$-polyregular
function} $f$ is \kl{commutative}, ensuring that we are working in a decidable
subclass of functions.

\begin{lemma}
    \label{decide-commutative:lemma}
    One can decide if a function $f \in \ZPoly$
    is \kl{commutative}.
\end{lemma}
\begin{proof}
    Let $\Sigma$ be a finite alphabet, endowed with a
    total ordering over its letters,
    and let $f \colon \Sigma^* \to \Rel$ be a
    \kl{$\Rel$-polyregular function}.
    The map $\polysort \colon \Sigma^* \to \Sigma^*$
    that sorts the letters in the input word $w$ according
    to the chosen ordering is a \kl{polyregular function}
    \cite[Proposition II.12]{LOPEZ23b}.
    As a consequence,
    $(f \circ \polysort) \in \ZPoly$.
    Finally, $f$ is \kl{commutative} if and only
    if $f = f \circ \polysort$, which is decidable
    in the case of $\ZPoly$
    \cite[Corollary II.24]{LOPEZ23b}, and more generally in the case 
    of $\ZRat$ \cite[Corollary 3.6 p 38]{berstel2011noncommutative}.
\end{proof}

\cref{decide-commutative:lemma} does not state decidability in the case of
\kl{$\Rel$-rational series}, and its proof relies on the stability of $\ZPoly$
under the pre-composition by \kl{polyregular functions}, which does not hold
for $\ZRat$. However, and for the sake of completeness, let us sketch a (more
involved) proof that \kl{commutativity} is decidable for this larger class,
using non-trivial results coming from topological things.

\begin{lemma}
    \label{decidable-commutative-rat:lemma}
    Let $f \in \ZRat$, it is decidable whether
    $f$ is \kl{commutative}.
\end{lemma}
\begin{proof}
    Because $f$ is a \kl{$\Rel$-rational series},
    there exists a $n \in \Nat$,
    vectors $v_1$ and $v_2$,
    and $n$ by $n$ integer matrices $\seqof{M_a}{a \in \Sigma}$,
    such that
    \begin{equation*}
        \forall w \in \Sigma^*,
        f(w) = {}^t v_1 \left(\prod_{i = 1}^{|w|} M_{w_i}\right) v_2 \quad .
    \end{equation*}
    To simplify notations, 
    let us write $M_w \defined \prod_{i = 1}^{|w|} M_{w_i}$,
    and 
    $S \defined \setof{M_w}{w \in \Sigma^+}$ be the
    semigroup generated by $\seqof{M_a}{a \in \Sigma}$.

    The function $f$ is \kl{commutative} if and only if for all $u,v \in
    \Sigma^+$, ${}^t v_1 (M_u M_v  - M_v M_u) v_2 = 0$. The above is a
    polynomial equation involving the coefficients of $M_u$ and $M_v$.
    Let us compute $\overline{S \times S}$ the Zariski closure of $S \times S$.
    The above polynomial vanishes over $S \times S$ if and only
    if it belongs to the Zariski closure, which is decidable
    \cite{HROUPOWO18}.
\end{proof}

### 2024-01-17 parts/starfree.tex
residual transducer??

@@

\begin{theorem}
    \label{decidable-n-poly:thm}
    Given a \kl{commutative}
    \kl{$\Rel$-polyregular function} $f \colon \Sigma^* \to \Rel$,
    one can decide if $f$ can be computed by a \kl{$\Nat$-polyregular function},
    and effectively compute a representation of $f$.
\end{theorem}

To prove the theorem, let us first provide a nice characterization of
\kl{commutative} \kl{$\Rel$-polyregular functions} in terms of polynomials.
This is actually a refinement of \cref{n-poly-combinatorics:lem} in the
commutative case.

\begin{lemma}
    \label{decompose-polynomial:lem}
    Let $f \colon \Sigma^* \to \Rel$ be a \kl{commutative}
    \kl{$\Rel$-polyregular function}. There exists $\omega \in \Nat$,
    and polynomials $\seqof{P_{\seqof{n_a}{a \in \Sigma}}}{ 0 \leq n_a < \omega}$,
    such that for all $w \in \Sigma^*$,
    \begin{equation*}
        f(w) = P_{\seqof{\card[a]{w} [\omega]}{a \in \Sigma}}
        (\seqof{\card[a]{w} / \omega}{a \in \Sigma}) \quad .
    \end{equation*}
    Which informally means that for every choice of rest modulo $\omega$
    for every variable, there exists a polynomial describing the
    behavior of $f$ when adding multiples of $\omega$ letters at a time.

    Furthermore, the polynomials are computable from $f$.
\end{lemma}
\begin{proof}
    equivalent partitions of the word, but commutativity?.q
    Lemma 5.37 phd gaëtan.
\end{proof}

@@

\begin{conjecture}
    $\CorrectPoly$ is precisely composed of polynomials that
    have non-negative coefficients in a \emph{binomial basis},
    i.e., where polynomials are obtained
    as sums and products of $\binom{X_i}{p_i}$.
\end{conjecture}

@@

Let us now prove that the above characterizations of
commutative
$\Nat$-polyregular functions can be combined with the recent
advances in the study of $\Rel$-polyregular functions 
\cite{LOPEZ23b}
to prove the following theorem.

\begin{theorem}
    \label{zsf-npoly-nsf:thm}
    Let $\Sigma$ be a finite alphabet, and $f \colon \Sigma^* \to \Nat$
    be both a \kl{star-free $\Rel$-polyregular function}
    and a \kl{commutative} \kl{$\Nat$-polyregular function}.
    Then, $f$ is a \kl{commutative} \kl{star-free $\Nat$-polyregular function}.
    That is, the following equality holds:
    \begin{equation*}
        \ZCommut \cap \ZSF[k] \cap \NPoly
        = \ZCommut \cap \ZSF \cap \NPoly[k]
        = \ZCommut \cap \NSF
        \quad .
    \end{equation*}
\end{theorem}

@@

\begin{conjecture}
    \label{nsf-zsf:conj}
    In the general case, the following equation holds:
    \begin{equation*}
        \ZSF \cap \NPoly
        = 
        \NSF
        \quad .
    \end{equation*}
\end{conjecture}

@@

\begin{corollary}
    One can decide if a \kl{commutative} \kl{ $\Rel$-polyregular function}
    can be realized by a \kl{star-free $\Nat$-polyregular function},
    in which case the conversion is effective.
\end{corollary}

@@

>>>>

\subsection{$\Nat$-polyregular functions}

@@

\begin{definition}
    \label{k-combinatorial:def}
    Let $k \in \Nat$, and $f \colon \Sigma^* \to \Rel$
    be a \kl{$\Rel$-polyregular function}. The function $f$ is 
    \intro{$(k,\Nat)$-combinatorial} if there exists $\omega \in \Nat$,
    such that
    for all
    $\alpha_0, \dots, \alpha_k \in \Sigma^*$,
    for all $u_1, \dots, u_k \in \Sigma^*$,
    there exists a \kl{$\Nat$-rational polynomial} $P$
    satisfying for all $X_1, \dots, X_k \geq \omega$:
    \begin{equation*}
        f
        \left(
            \alpha_0 \prod_{i = 1}^k u_i^{\omega X_i} \alpha_i
        \right)
        = 
        P(X_1, \dots, X_k) \quad .
    \end{equation*}
\end{definition}

\begin{theorem}
    \label{decidable-n-poly:thm}
    Let $k \in \Nat$, and 
    let $f \in \ZPoly[k]$ be a \kl{commutative} \kl{$\Rel$-polyregular function}.
    The following are equivalent:
    \begin{enumerate}
        \item $f$ is \kl{$(k,\Nat)$-combinatorial},
        \item $f \in \NPoly$,
    \end{enumerate}
    Furthermore, the properties are decidable,
    and one can effectively compute a representation of $f$.
\end{theorem}

To prove the theorem, let us first provide a nice characterization of
\kl{commutative} \kl{$\Rel$-polyregular functions} in terms of polynomials.
This is actually a refinement of \cref{n-poly-combinatorics:lem} in the
commutative case. Informally, we state that \kl{commutative}
\kl{$\Rel$-polyregular functions} are obtained by applying different
\kl{$\Rel$-rational polynomials} based on the modulos...

\AP Given $\omega \in \Nat$, let us write $\phi \colon \Nat \to \set{0, \dots,
\omega} \times \set{0, \dots, \omega - 1}$ defined as $\phi(n) = (t, r)$ where
$t = \min(n, \omega)$, and $r$ is the rest of the Euclidian division of $n$ by
$\omega$.
We let $\phi^\dagger \colon \Nat^k \to$.

\begin{lemma}
    \label{decompose-polynomial:lem}
    Let $f \colon \Sigma^* \to \Rel$ be a \kl{commutative}
    \kl{$\Rel$-polyregular function}. There exists a computable
    $\omega \in \Nat$,
    and computable 
    polynomials $\seqof{P_{\rho}}{\rho \in \Delta_\omega}$,
    such that for all $w \in \Sigma^*$,
    \begin{equation*}
        f\left(a_1^{X_1} \cdots a_n ^{X_n}\right) = P_{\phi(X_1, \dots, X_n)} (X_1 / \omega, \dots, X_n / \omega)
        \quad .
    \end{equation*}
\end{lemma}
\begin{proof}
    equivalent partitions of the word, but commutativity?.q
    Lemma 5.37 phd gaëtan.
\end{proof}

@@

@@

Let us now prove that the above characterizations of \kl{commutative}
\kl{$\Nat$-polyregular functions} can be combined with the recent advances in
the study of \kl{$\Rel$-polyregular functions} \cite{LOPEZ23b} to prove the
following theorem.

\begin{theorem}
    \label{zsf-npoly-nsf:thm}
    Let $\Sigma$ be a finite alphabet and $k \in \Nat$.
    Then,
    \begin{equation*}
        \ZCommut \cap \ZSF[k] \cap \NPoly
        = \ZCommut \cap \NSF[k]
        \quad .
    \end{equation*}
\end{theorem}

@@

\begin{conjecture}
    \label{nsf-zsf:conj}
    In the general case, the following equation holds for all $k \in \Nat$:
    \begin{equation*}
        \ZSF[k] \cap \NPoly
        = 
        \NSF[k]
        \quad .
    \end{equation*}
\end{conjecture}

@@

\textbf{TODO: add the growth rate here, we are actually minimising it}.
\begin{corollary}
    One can decide if a \kl{commutative} \kl{$\Rel$-polyregular function}
    can be realized by a \kl{star-free $\Nat$-polyregular function},
    in which case the conversion is effective.
\end{corollary}
\begin{proof}
    From \cite[Theorem V.13]{LOPEZ23b}, it is decidable whether
    $f \in \ZSF[k]$ and from \cref{decidable-n-poly:thm}
    it is decidable whether $f \in \NPoly$.
    Furthermore, one can enumerate functions in $\NSF[k]$,
    and check equivalence to effectively represent $f$ as a
    \kl{star-free $\Nat$-polyregular function}.
\end{proof}

@@

\subsection{Non-negative Residual Transducers}

The goal of the upcoming section is to provide a notion of residual transducer
for \kl{commutative} \kl{$\Nat$-polyregular functions}.

\begin{fact}
    For every $K \in \Nat$, and every $n \in \Nat$,
    $\Nat^n$ is a \kl{well-quasi-ordering}
    when endowed with the quasi-ordering $\vec{u} \leq_K \vec{v}$
    if and only if $\vec{u} = \vec{v}$ or
    $\vec{u}$ is smaller than $\vec{v}$ for the product ordering and
    $\sum_{i = 1}^n  (\vec{v} - \vec{u})_i \geq K$.
\end{fact}
\begin{proof}
    First, $\Nat^n$ with the product ordering is a \kl{well-quasi-ordering}.
    Therefore, given a sequence $\seqof{\vec{u}_i}{i \in \Nat}$,
    one can extract an infinite monotone subsequence for
    the product ordering.
    Either there exists $i < j$ such that $\vec{u}_i = \vec{u}_j$
    in the subsequence, in which case $\vec{u}_i \leq \vec{u}_j$
    and the original sequence was good. Or for all $i < j$,
    we have $\vec{u}_i$ that is strictly smaller than 
    $\vec{u}_j$ for the product ordering. As a consequence,
    $\vec{u}_0 \leq \vec{u}_K$.
\end{proof}

\begin{fact}
    Let $P \in \Nat[X,\vec{Y}]$, and $K,L \in \Nat$,
    then $P(X+L, \vec{Y}) - P(X+K,\vec{Y}) \in \Nat[X, \vec{Y}]$
    whenever $L \geq K$.
\end{fact}
\begin{proof}
    We can write
    $P = \sum_{i = 0}^n P_i(\vec{Y}) X^i$, with $P_i \in \Nat[\vec{Y}]$.
    Now,
    $P(X+L,\vec{Y}) - P(X+K,\vec{Y})$
    is precisely
    $\sum_{i = 1}^n P_i(\vec{Y}) (L-K) \left[ \sum_{j + l + 1= i} (X+L)^j (X+K)^l \right]$,
    remaining with positive coefficients.
\end{proof}

\begin{fact}
    Let $P \in \CorrectPoly[X,\vec{Y}]$, and $K,L \in \Nat$,
    then $P(X+L, \vec{Y}) - P(X+K,\vec{Y}) \in \CorrectPoly$
    whenever $L \geq K$.
\end{fact}
\begin{proof}
    We can write
    $P = \sum_{i = 0}^n P_i(\vec{Y}) X^i$, with $P_i \in \Nat[\vec{Y}]$.
    Now,
    $P(X+L,\vec{Y}) - P(X+K,\vec{Y})$
    is precisely
    $\sum_{i = 1}^n P_i(\vec{Y}) (L-K) \left[ \sum_{j + l + 1= i} (X+L)^j (X+K)^l \right]$.
    Now, $P_i(\vec{Y})$
    remaining with positive coefficients.
\end{proof}

\begin{fact}
    Let $P \in \CorrectPoly[X_1, \dots, X_n]$. 
    For all
    $\vec{u} \leq \vec{v} \in \Nat^n$,
    $Q(X_1, \dots, X_n)
    \defined 
    P(X_1 + v_1, \dots, X_n + v_n)
    -
    P(X_1 + u_1, \dots, X_n + u_n)$
    belongs to $\CorrectPoly[X_1, \dots, X_n]$.
\end{fact}
\begin{proof}
    Let us consider $K \in \Nat$ such that for all partial valuations
    $\nu \colon \vec{X} \topartial \Nat$,
    $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.

    Remark that
    \begin{align*}
        Q
        &= 
        P(X_1 + v_1, \dots, X_n + v_n)
        -
        P(X_1 + u_1, \dots, X_n + u_n) \\
        &=
        P(X_1 + v_1, X_2 + v_2 , \dots, X_n + v_n) - P(X_1 + u_1, X_2 + v_2, \dots, X_n + v_n) \\
        &+
        P(X_1 + u_1, X_2 + v_2, \dots, X_n + v_n) - P(X_1 + u_1, X_2 + u_2, \dots, X_n + v_n) \\
        &+ \dots
    \end{align*}

    Let $\nu \colon \vec{X} \topartial \Nat$
    be a partial valuation, let us show that
    $\translate{K}(\restr{Q}{\nu}) \in \Nat[\vec{X}]$.

    \begin{align*}
        \translate{K}(\restr{Q}{\nu})
        &= 
        \translate{K}(\restr{P(X_1 + v_1, \dots, X_n + v_n)}{\nu})
        -
        \translate{K}(\restr{P(X_1 + u_1, \dots, X_n + u_n)}{\nu})
    \end{align*}
\end{proof}

\begin{lemma}
    Let $k \in \Nat$,
    $f \in \NPoly[k]$,
    and $\seqof{u_i}{i \in \Nat}$ be a sequence of
    elements in $\Nat^n$. There exists $i < j$
    such that
    $f(u_j \cdot) - f(u_i \cdot) \in \NPoly[k-1]$.
\end{lemma}

\begin{defintion}
    Residual $\Nat$-transducer.
\end{definition}

\begin{lemma}
    Computable 
\end{lemma}

\begin{theorem}
    If and only if this one is counter-free.
\end{theorem}

Note that it is not needed for decidability, as we are able to use directly the
results on $\ZSF$, but this provides an independent proof, and highlights the
versatility of this notion of "residual transducer".
